\Large Generation of Mathematical Programming Representation for Discrete Event Simulation Models of Timed Petri Nets

### 2023-01-09 Answer to reviewers.tex

\textit{2. Detailed Comments}

\textit{The paper reads clearly. There are only very few typos (see at the end).}
\textbf{Answer: we have correct the typos.}

\textit{From a technical point of view, there is one thing that needs to be clarified, page 8: How do you choose constant M? Understandably, it should be large enough. However, I would expect to find in the paper a discussion about a lower bound on M.}

\noindent
\textbf{Answer: In the subsections of Section 3, the value of $M$ is discussed each time it %appears
is used, as different values of $M$ are required in the constraints where $M$ is needed.}

\textbf{Section 3.1: The value of $M$ in constraints (A1) and (A2) can be set to a value that is larger than the simulation time span as it is needed to bound the finishing time of transitions.}

\textbf{Section 3.2: Also in this case, the value of $M$ can be set to a value that is larger than the simulation time span for constraints (B1) and (B2) as 
it is used to bound the transition firing time. %The value of $M$ can be chosen as the upper bound of marking of place $p$ minus $(W^{p,t}-1)$ for constraints (B4). 
For constraints (B4), instead, the value of $M$ can be chosen as the upper bound of marking of place $p$ minus $(W^{p,t}-1)$.}

\textbf{Section 3.3: The value of $M$ in constraints (C3) and (C4) can be set to $K$, i.e., the total number of iterations. }

### 2023-01-09 main.tex
Arianna

\bigskip
\begin{abstract}
This work proposes a mathematical programming (MP) representation of discrete event simulation of timed Petri nets (TPN). Currently, mathematical programming techniques are not widely applied to optimize discrete event systems due to the difficulty of formulating models capable to %catch 
correctly represent the system dynamics. This work connects the two fruitful research fields, mathematical programming and timed Petri nets. In the MP formalism, the decision variables of the model %are composed of 
correspond to the transition firing times and the markings of the TPN, whereas the constraints represent the state transition logic and temporal sequences among events. The MP model and a simulation run of the TPN are then totally equivalent, which is validated %with several cases. 
through several applications. Using a TPN model as input, the MP model can be routinely generated and used as white box for further analysis such as sensitivity analysis, cut generation in optimization procedures, and proof of formal properties.

@@

Discrete event dynamic systems (DEVS) have wide application in manufacturing and service applications. Even though many theoretical and analytical studies on discrete event dynamic systems have been developed, performance of real systems with high degree of complexity can, sometimes, only be evaluated with discrete event simulation (DES). Hence, simulation--optimization algorithms are widely used when numerical performance evaluation must be coupled with optimization, i.e., when the best system configuration, according to some criteria, has to be found meanwhile guaranteeing a given value of some performance measures \citep{fu20115book}.

@@

The rest of the paper is organized as follows. Section \ref{sec:sim} briefly introduces the TPN and its simulation algorithm. Section \ref{sec:TPN_MPR} describes the generation of the MP model from an existing TPN model. Section \ref{sec:example} shows examples of application and validation of the equivalence between simulation and the MP. Discussion and conclusion are reported in Section\ref{sec:conclude}.

@@

The illustrative example showed in Figure \ref{fig:gg1_petri} is a G/G/2 queue, in which customers arrive at the queue following an arbitrary arrival process. One of the two identical servers (or the idle one if only one is idle) processes customers in the queue one by one, and the processing time is arbitrarily distributed. After processing, the customer will be immediately released from the system, and the server becomes idle. At the right side of Figure \ref{fig:gg1_petri}, the TPN graph of the G/G/2 system is %presented. 
reported. A token in place $p_1$ shows that it is possible to schedule the arrival of the next customer, and the firing time of transition $t_1$ is equal to the inter-arrival time. At the finish of firing $t_1$, a customer arrives, and it is possible to schedule the next arrival, thus, it releases one token to $p_1$. Furthermore, an arrival will increase the length of the queue by 1, which is represented by the number of tokens in place $p_2$. Two tokens in place $p_3$ shows that two servers are idle. The firing of transition $t_2$ shows the processing of a customer, and it is enabled if and only if at least one server is idle and there is a customer in the queue. Once the process of the customer finishes, i.e., the finish of firing $t_2$, a server is released, and a token is sent to place $p_3$.

@@

First of all, sets $\mathbb{K}$, $\mathbb{T}$ and $\mathbb{I}^t$ are used to indicate the set of iterations, the set of transitions and the set of firings of transition $t$, respectively.
Binary variables $w^{t}_{i,k}$ are introduced to represent that the $i$-th start firing of transition $t$ finishes in iteration $k$. If $w^{t}_{i,k}$  is equal to one, %and 
variable $e^{t,f}_{i}$ is bounded to $\mathcal{E}_k$, as shown in constraints (A1) and (A2). Constraints (A3) state that each firing-finish event executes at most once. Constraints (A4) state that, in each iteration, exactly one firing-finish event is executed. Constraints (A5) implies that the delay between the start and finish of firing transition $t$ is equal to a sample from the random variable $T^{t}$. Constraints (A6) imply that the clock cannot be reversed from iteration $k-1$ to iteration $k$. Constraint (A7) implies that the simulation clock is initialized to zero. The value of $M$ in constraints (A1) and (A2) can be set to a value that is larger than the simulation time span.

@@

The second group of constraints, denoted by group--B, depicts the start of transition firing, as in lines 5 to 13 in Algorithm 1. Binary variables $x^{t}_{i,k}$ represent that a fire of transition $t$ starts in iteration $k$ if it is equal to one. Constraints (B1) and (B2) shows that the start firing time is equal to the simulation clock. The value of $M$ can be set to a value that is larger than the simulation time span.

@@

\subsection{Event sequencing}
The index $i$ of event $e^{t,s}_i$ represents the sequence of firing starting, i.e., if a firing starts in an earlier iteration, then its index will be smaller. Moreover, a firing must finish after starting. Group--C constraints force such sequences, which are relevant to all the events. Constraints (C1) show that if the $i$--th firing of transition $t$ does not start before simulation termination, then the $(i+1)$--th firing will not start. Constraints (C2) state that if the $i$--th firing of transition $t$ does not start before simulation termination, then it will not finish. Constraints (C3) state that a firing cannot finish before starting, unless it remains in the future event list at the end of the simulation run. Constraints (C4) depict that the $(i+1)$--th firing of transition $t$ must be scheduled after the $i$--th execution, unless the $(i+1)$--th execution is not scheduled before simulation termination. The value of $M$ in constraints (C3) and (C4) can be set to $K$.

@@

\section{Indicative examples} \label{sec:example}
In this section, the $G/G/m$ is used as a example to generate MPR from TPN. Both the proposed and the state-of-the-art formulation are both presented. Finally, an extended model to optimize the $G/G/m$ queue is stated.

\subsection{MPR of G/G/m queue}\label{sec:MPR_ggm}
A $G/G/m$ queue is composed by $m$ parallel identical servers. Customers arrive at the queue following an general arrival process. One of the identical idle servers processes customers in the queue one by one, and the processing time is generally distributed. After being processed, the customer will be released from the system immediately, and the server becomes idle again. In Figure \ref{fig:tpn_GG2}, the TPN of $G/G/2$ is presented. A token in place $p_{arr}$ shows that it is possible to schedule the arrival of next customer, and the firing time of transition $t_{arr}$ is equal to the inter-arrival time. At the end of firing $t_{arr}$, a customer arrives, and it is possible to schedule the next arrival, thus, it releases one token to $p_{arr}$. Furthermore, an arrival will increase the length of the queue by 1, which is represented by the number of tokens in place $p_{queue}$. Two tokens in place $p_{idle}$ shows that two servers are idle. The firing of transition $t_{process}$ represents the processing of a customer, and it is enabled if and only if at least one server is idle and there is a customer in the queue. Once the process of the customer finishes, i.e., the end of firing $t_{process}$, a server is released, and a token is sent to place $p_{idle}$.

@@

The sets $\mathbb{T},\ \mathbb{K},\ \mathbb{I}^t$ are specified as follows. Set $\mathbb{T}$ is composed of three transitions $\{t_{arr},\tilde{t}_{process}, \bar{t}_{process}\}$ as shown in Figure \ref{fig:tpn_GG2_expansion}. If $n$ jobs' arrival, processing and departure are simulated, the three transitions are fired for $n$ times each, and set $\mathbb{I}^t$ is equal to $\{1,2,..., n\}$. There are $3n$ transition firings in the simulation execution and the set $\mathbb{K}$ is equal to $\{0,1,..., 3n-1\}$. The initial markings $m^{p_{arr}}_0,\ m^{p_{queue}}_0, m^{p_{process}}_0, m^{p_{idle}}$ are specified as 1, 0, 0, 2. Once the TPN with the initial markings and simulation length are provided, the MPR composed of (A1) to (A7), (B1) to (B7), (C1) to (C4) and (D1) can be generated.

@@

Table \ref{tab:sim_GGM} shows the first ten iteration of a simulation run of a G/G/m queue.

In iteration 0, the system is in its initial state with markings (1,0,0,2), and transition $t_{arr}$ can be fired, equivalently, binary variables $z^{arr}_0$ and $x^{arr}_{1,0}$ are equal to $1$ in the solution of the MPR. The time to start fire transition $t_{arr}$ is equal to $0$, and the variable $e^{arr,s}_{1}$ takes value $0$ in the MPR solution. The fire will finish at time 2.3, and variable $e^{arr,f}_{1}$ takes value 2.3 equivalently. That firing is also the firing to finish in this iteration, and $w^{arr}_{1,0}$ is equal to 1. After that, the markings are changed to (1,1,0,2) and this iteration ends.

At the beginning of iteration 1, the simulation clock is advanced to the time of the firing finished in iteration 0, i.e., $2.3$. In the MPR solution, the variable $\mathcal{E}_1$ is equal to $2.3$. With markings (1,1,0,2), it is possible to fire $t^{arr}$ and $\tilde{t}^{proc}$, and variables $z_1^{arr}$ and $z_1^{\tilde{proc}}$ are equal to 1. This firings are the second and first firing of $t^{arr}$ and $\tilde{t}^{proc}$, respectively, and variables $x_{2,1}^{arr}$ and $x_{1,1}^{\tilde{proc}}$ take value 1 in the MPR solution. The start time of the transitions $t^{arr}$ and $\tilde{t}^{proc}$ are equal to the clock time, and variables $e^{arr,s}_{2}$ and $e^{\tilde{proc},s}_{1}$ are equal to $2.3$. With a sample from the arrival process, next customer will arrive at time 11.1, i.e., the finish time of transition $t^{arr}$ is equal to 11.1, which is the value of variable $e^{arr,f}_{2}$. Since transition $\tilde{t}^{proc}$ is zero-firing-time, the firing will finish at time 2.3, which is the value of variable $e^{\tilde{proc},f}_{1}$. There are two firing to finish in the queue, and the one of $\tilde{t}^{proc}$ has earlier time, so it will be the firing to finish in this iteration and variable $w^{\tilde{proc}}_{1,1}$ is equal to one. After that, the markings are changed to $(0,0,1,1)$ and this iteration ends.

Similarly for iterations 2 to 10, the values of the MPR variables in the solution and the simulation realization can be justified as shown in the Table \ref{tab:sim_GGM}.

@@

\subsection{Comparison with Schruben's model of G/G/m systems}
In \cite{chan2008optimization}, an MPR of DEVS simulation based on ERG is presented. The ERG of the $G/G/2$ queue is shown in Figure \ref{fig:ERG_GG2}. In the ERG, nodes and arcs represent the events and the triggering relationships. The simulation of the $G/G/2$ queue is composed of three events, which are $arrival$, $start\ process$ and $finish\ process$ respectively. The three events are presented by nodes $Arr$, $Start\ proc$ and $Fin\ proc$ in the ERG. The node $Run$ indicates the launch of the simulation. There are two state variables for $G/G/2$ queue, $queue$ and $idle$, representing the number of jobs in the queue and the number of idle servers. The equations below each node state how the state variables are changed upon the execution of the events. For instance, when event $Arr$ occurs, one job enters the system and $queue$ is incremented by one. Where there is an arc connecting two nodes, it is possible to execute the destination event after the source event execution under certain conditions. The condition can be a time delay or a logic expression, labeled on the arcs. The arc pointing from $Arr$ to $Arr$ is accompanied with a time delay $\tau{arr}$, indicating that the next job will arrive with a time delay equal to the inter-arrival time after the previous one. The arc pointing from $Arr$ to $Start\ proc$ is accompanied with a logic expression $idle\ge 1$, indicating that after a job's arrival, it can start to be processed only if there is at least one idle server.

@@

The objective function \eqref{erg_1} is to minimize the execution time of all the events. Constraints \eqref{erg_2} to \eqref{erg_8} are derived from the arcs in the ERG, indicating the event triggering relationships. The arc pointing from $Arr$ to $Arr$ generates the constraints \eqref{erg_2}, and shows the inter-arrival time between two adjacent arrivals. The arc pointing from $Arr$ to $Start\ proc$ generates the constraints \eqref{erg_3}, and shows that a job is processed after it arrives. The arc pointing from $Start\ proc$ to $Fin\ proc$ generates the constraints \eqref{erg_4} to \eqref{erg_7}, and shows that a job is finished after the process starts with certain processing time. Since the processing time is generated from probability distributions, it can be unequal for different jobs. A job starts later can surpass an early one, i.e., the order of the $i$-th start job in the finish sequence can be different from $i$. Therefore, binary variables $\delta^{Start\ proc, Fin\ proc}_{i,i^{'}}$ are introduced, and the $i$-th start job is the $i^{'}$-th to finish if and only if $\delta^{Start\ proc, Fin\ proc}_{i,i^{'}}$ is equal to one. The arc pointing from $Fin\ proc$ to $Start\ proc$ generates the constraints \eqref{erg_8}, and shows that a job can be started when there is an idle server. The subscripts $i$ of variables $e^{Start\ proc}_{i}$ and $e^{Fin\ proc}_{i}$ indicates the occurrence sequence of the events, as in constraints \eqref{erg_9} and \eqref{erg_10}.

This MPR model is different from the one proposed in this work in the following aspects. First, the system states are not explicitly modeled in this model, which are decision variables in the proposed model. Second, the objective is to minimize all the event execution time in this model. However, the formulation of objective function can be rather flexible in the proposed model, since executing events as soon as possible is assured by the constraints. Third, \cite{chan2008optimization} only proposed the descriptive procedure to generate an MPR, without well-formulated mapping from ERG to MPR, which makes the application difficult. In this work, the proposed method is well-formulated, and the input required is only the timed Petri net.

\subsection{MPR optimizing server number of G/G/m queue}
An MP model optimizing the server number in a G/G/m queue based on the MPR proposed in Section \ref{sec:MPR_ggm} is as follows. The objective function \eqref{opt_1} is to minimize the number of servers, which is the initial marking in place $p^{idle}$. Constraint \eqref{opt_2} indicates that the average waiting time does not exceed a target $wt$. Constraints (A1) - (A7), (B1) - (B7), (C1) - (C4), (D1) are the same as presented in Section \ref{sec:MPR_ggm}.

@@

\section{Conclusion} \label{sec:conclude}
This work proposes a well-formulated framework to translate general TPNs into MP models. With an example of $G/G/m$ queue, the convenience of the formulation procedure is demonstrated.

This work lays a common foundation for developing MP--based approaches that can be used to analyze and optimize TPNs, based on the vast literature available in the mathematical programming field. Even though many existing works in literature are related to this research direction on specific applications, especially in manufacturing systems, a methodological work is still a blank, which will be the future development.

>>>>

\bigskip
\begin{abstract}
This work proposes a mathematical programming (MP) representation of discrete event simulation of timed Petri nets (TPN). Currently, mathematical programming techniques are not widely applied to optimize discrete event systems due to the difficulty of formulating models capable to %catch 
correctly represent the system dynamics. This work connects the two fruitful research fields, mathematical programming and timed Petri nets. In the MP formalism, the decision variables of the model %are composed of 
correspond to the transition firing times and the markings of the TPN, whereas the constraints represent the state transition logic and temporal sequences among events. The MP model and a simulation run of the TPN are then totally equivalent, which is validated %with several cases. 
through several applications. Using a TPN model as input, the MP model can be routinely generated and used as white box for further %analysis 
tasks such as sensitivity analysis, cut generation in optimization procedures, and proof of formal properties.

@@

Discrete event dynamic systems (DEVS) have wide application in manufacturing and service applications. Even though many theoretical and analytical studies on discrete event dynamic systems have been developed, performance of real systems with high degree of complexity can be, sometimes, only evaluated with discrete event simulation (DES). Hence, simulation--optimization algorithms are widely used when numerical performance evaluation must be coupled with optimization, i.e., when the best system configuration, according to some criteria, has to be found meanwhile guaranteeing a given value of some performance measures \citep{fu20115book}.

@@

The rest of the paper is organized as follows. Section \ref{sec:sim} briefly introduces the TPN and its simulation algorithm. Section \ref{sec:TPN_MPR} describes the generation of the MP model from an existing TPN model. Section \ref{sec:example} shows examples of application and validation of the equivalence between simulation and the MP. Discussion and conclusion are reported in Section \ref{sec:conclude}.

@@

The illustrative example showed in Figure \ref{fig:gg1_petri} is a G/G/2 queue, in which customers arrive at the queue following an arbitrary arrival process. One of the two identical servers (or the idle one if only one is idle) processes customers in the queue one by one, and the processing time is arbitrarily distributed. After processing, the customer will be immediately released from the system, and the server becomes idle. On the right side of Figure \ref{fig:gg1_petri}, the TPN graph of the G/G/2 system is %presented. 
reported. A token in place $p_1$ shows that it is possible to schedule the arrival of the next customer, and the firing time of transition $t_1$ is equal to the inter-arrival time. At the finish of firing $t_1$, a customer arrives, and it is possible to schedule the next arrival, thus, it releases one token to $p_1$. Furthermore, an arrival will increase the length of the queue by 1, which is represented by the number of tokens in place $p_2$. Two tokens in place $p_3$ shows that two servers are idle. The firing of transition $t_2$ shows the processing of a customer, and it is enabled if and only if at least one server is idle and there is a customer in the queue. Once the process of the customer finishes, i.e., the finish of firing $t_2$, a server is released, and a token is sent to place $p_3$.

@@

First of all, sets $\mathbb{K}$, $\mathbb{T}$ and $\mathbb{I}^t$ are used to indicate the set of iterations, the set of transitions and the set of firings of transition $t$, respectively.
Binary variables $w^{t}_{i,k}$ are introduced to represent that the $i$-th start firing of transition $t$ finishes in iteration $k$. If $w^{t}_{i,k}$  is equal to one, %and 
variable $e^{t,f}_{i}$ is bounded to $\mathcal{E}_k$, as shown in constraints (A1) and (A2). Constraints (A3) state that each firing-finish event executes at most once. Constraints (A4) state that, in each iteration, exactly one firing-finish event is executed. Constraints (A5) implies that the delay between the start and finish of firing transition $t$ is equal to a sample from the random variable $T^{t}$. Constraints (A6) imply that the clock cannot be reversed from iteration $k-1$ to iteration $k$. Constraint (A7) implies that the simulation clock is initialized to zero. The value of $M$ in constraints (A1) and (A2) can be set to a value that is larger than the simulation time span, as it is needed to bound the finishing time of transitions.

@@

The second group of constraints, denoted by group--B, depicts the start of transition firing, as in lines 5 to 13 in Algorithm 1. Binary variables $x^{t}_{i,k}$ represent that a fire of transition $t$ starts in iteration $k$ if it is equal to one. Constraints (B1) and (B2) shows that the start firing time is equal to the simulation clock. Also in this case, the value of $M$ can be set to a value that is larger than the simulation time span as it is used to bound the transition firing time.

@@

\subsection{Event sequencing}
The index $i$ of event $e^{t,s}_i$ represents the sequence of firing starting, i.e., if a firing starts in an earlier iteration, then its index will be smaller. Moreover, a firing must finish after starting. Group--C constraints force such sequences, which are relevant to all the events. Constraints (C1) show that if the $i$--th firing of transition $t$ does not start before simulation termination, then the $(i+1)$--th firing will not start. Constraints (C2) state that if the $i$--th firing of transition $t$ does not start before simulation termination, then it will not finish. Constraints (C3) state that a firing cannot finish before starting, unless it remains in the future event list at the end of the simulation run. Constraints (C4) depict that the $(i+1)$--th firing of transition $t$ must be scheduled after the $i$--th execution, unless the $(i+1)$--th execution is not scheduled before simulation termination. The value of $M$ in constraints (C3) and (C4) can be set to $K$, i.e., the total number of iterations.

@@

%\section{Indicative examples} 
\section{Examples of MPR generation from TPN}
\label{sec:example}
In this section, the $G/G/m$ is used as a example to generate MPR from TPN. Both the proposed and the state-of-the-art formulation are presented. Finally, also an extended model to optimize the $G/G/m$ queue is stated.

\subsection{MPR of G/G/m queue}\label{sec:MPR_ggm}
A $G/G/m$ queue is composed by $m$ parallel identical servers. Customers arrive at the queue following an general arrival process. One of the identical idle servers processes customers in the queue one by one, and the processing time  of each customer is generally distributed. After being processed, the customer will be released from the system immediately, and the server becomes idle again. In Figure \ref{fig:tpn_GG2}, the TPN of $G/G/2$ is shown. A token in place $p_{arr}$ shows that it is possible to schedule the arrival of the next customer, and the firing time of transition $t_{arr}$ is equal to the inter-arrival time. At the end of firing of $t_{arr}$, a customer arrives, and it is  then possible to schedule the next arrival, i.e., it releases one token to $p_{arr}$. Furthermore, each arrival will increase the length of the queue by 1, which is represented by the number of tokens in place $p_{queue}$. Two tokens in place $p_{idle}$ shows that two servers are idle. The firing of transition $t_{process}$ represents the processing of a customer, and it is enabled if and only if at least one server is idle and there is at least one customer in the queue. Once the process of the customer finishes, i.e., the end of firing of $t_{process}$, the server is released, and a token is sent to place $p_{idle}$.

@@

The sets $\mathbb{T},\ \mathbb{K},\ \mathbb{I}^t$ are specified as follows. Set $\mathbb{T}$ is composed of three transitions $\{t_{arr},\tilde{t}_{process}, \bar{t}_{process}\}$ as shown in Figure \ref{fig:tpn_GG2_expansion}. If $n$ job arrival, processing and departure are simulated, each of the three transitions is fired for $n$ times, and the set $\mathbb{I}^t$ is equal to $\{1,2,..., n\}$. There are $3n$ transition firings in the simulation execution and the set $\mathbb{K}$ is equal to $\{0,1,..., 3n-1\}$. The initial markings $m^{p_{arr}}_0,\ m^{p_{queue}}_0, m^{p_{process}}_0, m^{p_{idle}}$ are specified as 1, 0, 0, 2. Once the TPN with the initial markings and the simulation length are provided, the MPR %composed of 
corresponding to constraints (A1) to (A7), (B1) to (B7), (C1) to (C4) and (D1) can be generated.

@@

\noindent
Table \ref{tab:sim_GGM} shows the first ten iteration of a simulation run of a G/G/m queue.

In iteration 0, the system is in its initial state with markings (1,0,0,2), and transition $t_{arr}$ can be fired; equivalently, binary variables $z^{arr}_0$ and $x^{arr}_{1,0}$ are equal to $1$ in the solution of the MPR. The time to start firing transition $t_{arr}$ is equal to $0$, and the variable $e^{arr,s}_{1}$ takes value $0$ in the MPR solution. The fire will finish at time 2.3, and variable $e^{arr,f}_{1}$ takes value 2.3 equivalently. %That firing is also the firing to finish in this iteration, and 
As that firing finished in the current iteration, $w^{arr}_{1,0}$ is equal to 1. After that, the markings are changed to (1,1,0,2) and the current iteration ends.

At the beginning of iteration 1, the simulation clock is advanced to the time of the firing has finished in iteration 0, i.e., $2.3$. In the MPR solution, the variable $\mathcal{E}_1$ is then equal to $2.3$. With markings (1,1,0,2), it is possible to fire $t^{arr}$ and $\tilde{t}^{proc}$, and variables $z_1^{arr}$ and $z_1^{\tilde{proc}}$ are equal to 1. These firings are the second and first firing of $t^{arr}$ and $\tilde{t}^{proc}$, respectively, and variables $x_{2,1}^{arr}$ and $x_{1,1}^{\tilde{proc}}$ take value 1 in the MPR solution. The starting time of the transitions $t^{arr}$ and $\tilde{t}^{proc}$ are equal to the clock time, and variables $e^{arr,s}_{2}$ and $e^{\tilde{proc},s}_{1}$ are equal to $2.3$. With a sample from the arrival process, the next customer will arrive at time 11.1, i.e., the finishing time of transition $t^{arr}$ is equal to 11.1, which is the value of variable $e^{arr,f}_{2}$. Since transition $\tilde{t}^{proc}$ is zero-firing-time, the firing will finish at time 2.3, which is the value of variable $e^{\tilde{proc},f}_{1}$. There are two firings to be finished in the queue, and that of $\tilde{t}^{proc}$ has earlier time, so it will be the firing that will be finished in this iteration and, hence, variable $w^{\tilde{proc}}_{1,1}$ is equal to one. After that, the markings are changed to $(0,0,1,1)$ and the iteration ends.

Similarly for iterations 2 to 10, the values of the MPR variables in the solution and the simulation realizations, %can be justified 
which are reported in Table \ref{tab:sim_GGM}, can be explained as done for iterations 0 and 1.

@@

\subsection{Comparison with Schruben's model of G/G/m systems}
In \cite{chan2008optimization}, an MPR of DEVS simulation based on ERG is presented. The ERG of the $G/G/2$ queue is shown in Figure \ref{fig:ERG_GG2}. In the ERG, nodes and arcs represent the events and the triggering relationships. The simulation of the $G/G/2$ queue is composed of three events, which are $arrival$, $starting\ process$ and $finishing\ process$, respectively. The three events are presented by nodes $Arr$, $Start\ proc$ and $Fin\ proc$ in the ERG. The node $Run$ indicates the launch of the simulation. There are two state variables for $G/G/2$ queue, $queue$ and $idle$, representing the number of jobs in the queue and the number of idle servers. The equations below each node state how the state variables are changed upon the execution of the events. For instance, when event $Arr$ occurs, one job enters the system and $queue$ is incremented by one. When there is an arc connecting two nodes, it means that it is possible to execute the destination event only after the source event is executed. %under certain conditions. 
The execution of an event can also be subject to certain conditions, reported as labels on the connecting arc. The conditions can be a time delay or a logic expression. %, labeled on the arcs. 
The arc pointing from $Arr$ to $Arr$ is accompanied with a time delay $\tau{arr}$, indicating that the next job will arrive with a time delay equal to the inter-arrival time from the previous one. The arc pointing from $Arr$ to $Start\ proc$ is %accompanied 
labeled with the logic expression $idle\ge 1$, indicating that after a job is arrived, it can start to be processed only if there is at least one idle server.

@@

The objective function \eqref{erg_1} is the minimization of the execution time of all the events. Constraints \eqref{erg_2} to \eqref{erg_8} are derived from the arcs in the ERG, and indicate the event triggering relationships. The arc pointing from $Arr$ to $Arr$ generates constraints \eqref{erg_2}, which represent the inter-arrival time between two adjacent arrivals. The arc pointing from $Arr$ to $Start\ proc$ generates  constraints \eqref{erg_3}, representing that a job can be processed only after it arrives. The arc pointing from $Start\ proc$ to $Fin\ proc$ generates the constraints \eqref{erg_4} to \eqref{erg_7}, representing that a job can be finished only after the process starts and a certain processing time is elapsed. Since the processing time is generated from probability distributions, it can take different values for different jobs. A job started later can be finished before an earlier stated one, i.e., the ranking of the $i$-th started job in the finishing sequence can be different from $i$. Therefore, binary variables $\delta^{Start\ proc, Fin\ proc}_{i,i^{'}}$ are introduced, and the $i$-th started job is the $i^{'}$-th to finish if and only if $\delta^{Start\ proc, Fin\ proc}_{i,i^{'}}$ is equal to one. The arc pointing from $Fin\ proc$ to $Start\ proc$ generates the constraints \eqref{erg_8}, and represent that a job can be started when there is an idle server. The subscripts $i$ of variables $e^{Start\ proc}_{i}$ and $e^{Fin\ proc}_{i}$ indicates the occurrence sequence of the events, as in constraints \eqref{erg_9} and \eqref{erg_10}.

This MPR model is different from the one proposed in this work in the following aspects. First, the system states, which are the decision variables in the model proposed in this work, are not explicitly modeled in the Chan and Schruben's model. Second, in the Chan and Schruben's model, the objective is to minimize all the event execution time. Instead, the formulation of objective function can be more flexible in the model proposed in this work, since executing events as soon as possible is already assured by the constraints. Third, \cite{chan2008optimization} only proposed the descriptive procedure to generate an MPR, without well-formulated mapping from ERG to MPR, which makes the application difficult. In this work, instead, the proposed method is well-formulated, and the only required input is the timed Petri net.

\subsection{MPR optimizing the server number of G/G/m queue}
An MP model optimizing the server number in a G/G/m queue, based on the MPR proposed in Section \ref{sec:MPR_ggm}, can be formulated as follows. 
As known, the larger the number of servers, the smaller the waiting time. However, servers have a cost, and this is why their number should be minimized.  
The objective function \eqref{opt_1} is then the minimization of the number of servers, which is the initial marking in place $p^{idle}$, i.e., variable $m^{idle}_0$. However, a constraints must be set on the maximum allowed waiting time, otherwise the obvious solution of number of servers equal to 1 is the one reached by the MPR solution.

@@

Constraint \eqref{opt_2} indicates that the average waiting time does not have to exceed a target value $wt$. Constraints (A1) - (A7), (B1) - (B7), (C1) - (C4), (D1) are the same as presented in Section \ref{sec:MPR_ggm}.

\section{Conclusion} \label{sec:conclude}
This work proposes a well-formulated framework to translate general TPNs into MP models. The convenience of the formulation procedure is demonstrated using the example of a $G/G/m$ queue.

This work lays a common foundation for developing MP--based approaches that can be used to analyze and optimize TPNs, based on the vast literature available in the mathematical programming field.

In fact, even though many existing works in the literature are related to this research direction, they are mainly focused on specific applications, especially in manufacturing systems, and a methodological work is still %a blank,
mussing, which will be the object of future research development.

\textbf{THE PAPER FINISHES HERE?}
 %%%%%%%%%%%%%%%%%%%%%%%%%
